The Resurgence of the Linear Optics Interferometer --- Recent Advances \& Applications

### 2017-05-04 paper.tex
Quantum metrology
Added blurb about NOON states and Heisenberg-limited interferometry.
Also described somewhat the MORDOR scheme. More ending comments needed.

\subsubsection{Non-PNR Detectors}
The first type of non-PNR detectors are photomultiplier tubes (PMTs). A PMT comprises of a vacuum tube with a photocathode for light absorption. The photoelectric effect then releases electrons in the photocathode, and this current is multiplied by a cascade of secondary electron emissions from a series of electrodes. This amplified signal can be picked up to indicate the presence of photons.

@@

\subsubsection{PNR detectors}

@@

  \subsection{Boson-sampling}

\subsection{Quantum metrology}

\comment{Discuss NOON states - Heisenberg limited}

\comment{Discuss MORDOR scheme}

>>>>

\subsubsection{Non-photon-number-resolving detectors}
The first type of non-PNR detectors are photomultiplier tubes (PMTs). A PMT comprises of a vacuum tube with a photocathode for light absorption. The photoelectric effect then releases electrons in the photocathode, and this current is multiplied by a cascade of secondary electron emissions from a series of electrodes. This amplified signal can be picked up to indicate the presence of photons.

@@

\subsubsection{Photon-number-resolving detectors}

@@

\subsection{Boson-sampling}\label{sec:bosonsampling}
\sihui{Peter to do this.}

\subsection{Quantum metrology}
An archetypal task for quantum metrology is the following: Given a Mach-Zedner interferometer, with an unknown phase shift of magnitude $\varphi$ inserted in one of the two paths (see Fig. \ref{fig:MZ}), can one deduce $\varphi$ via a judicious use of quantum states and measurements with a higher sensitivity than that of a completely classical interferometer? It is well-known that one indeed can, and the standard deviation, $\Delta\varphi$, after $N$ trials using quantum estimation is
\begin{align}
\Delta\varphi=\frac{1}{N} \ ,	
\end{align}
which gets a $\sqrt{N}$ improvement over classical phase estimation. As this sensitivity is also a fundamental limit arising from Heisenberg's uncertainty principle; we cannot measure definitively the phase, it is known as the Heisenberg limit. Most famously, a family of entangled states known as NOON states achieve this limit using $N$ photons in a single trial. A NOON state is a superposition state in which $N$ photons are either in the first or second path:
\begin{align}
\ket{\psi^{\rm (NOON)}_n}=\frac{1}{\sqrt{2}}(\ket{N,0}+\ket{0,N})\ .	
\end{align}
 Many reviews have been written for quantum metrology and Heisenberg-limited quantum phase estimation \cite{bib:Giovannetti04,bib:Dowling08,bib:Giovannetti11} so we will not dwell on the history of this application. Instead, we will discuss a new paradigm for quantum metrology that has arisen from our understanding of Boson-sampling.

 \begin{figure}[!htb]
\includegraphics[width=0.7\columnwidth]{phE}
\caption{A schematic diagram of the Mach-Zehnder interferometer comprising of two 50/50 beamsplitters with a phase shift $\varphi$ in one of its two paths.} \label{fig:MZ}	
\end{figure}

 As discussed in Section \ref{sec:bosonsampling}, the resources needed for Boson-sampling is readily available, but executes a task that is believed to be hard classically. Surprisingly, an entanglement between a large number of spatial modes is possible with single photons in a passive linear optical device. This so-called number-path entanglement can be harnessed for quantum metrology without the use of entangled states like the NOON state \cite{bib:Motes15}. Here, the phase shift to be estimated, $\varphi$, manifests itself as a linear phase gradient across $N$ modes (See Fig. \ref{fig:MORDOR}). By straddling these modes among a quantum Fourier transform and its conjugate operation, it is possible to achieve 
\begin{align}
\Delta \varphi=\mathcal{O}\left(\frac{1}{N^{3/2}}\right) \ ,	
\end{align}
with just a string of $N$ single photons, one photon per mode, for up to at least $N=25$ modes.

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{MORDOR}
\caption{Architecture of the quantum Fourier transform interferometer for metrology using single-photon states. QFT: Quantum Fourier transform circuit, and $\varphi_k=(k-1)\varphi$. Single photons are injected into the input modes on the left-hand-side, and a $N$-fold coincidence measurement is conducted at the output on the right-hand-side.} \label{fig:MORDOR}	
\end{figure}
